Faster generic IND-CCA2 secure KEM using ``encrypt-then-MAC''

### 2024-09-30 etm-transform/manuscript/main.tex
Checkpoint

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks. Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{nist2024module}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.

The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems. One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} \cite{bernstein2018towards} for providing ciphertext integrity. In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine. While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.

>>>>

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
The Fujisaki-Okamoto transformation \cite{DBLP:conf/crypto/FujisakiO99} is a generic construction that takes cryptographic primitives of lesser security and constructs a public-key encryption scheme with indistinguishability under adaptive chosen ciphertext attacks. Later works \cite{hofheinz2017modular}\cite{DBLP:conf/asiacrypt/HovelmannsHM22} 
\cite{bernstein2018towards} extended the original transformation to the construction of key encapsulation mechanism, which has been adopted by many post-quantum schemes such as Kyber \cite{bos2018crystals} (later standardized into ML-KEM \cite{FIPS203}), FrodoKEM \cite{bos2016frodo}, and SABER \cite{d2018saber}.

The current state of the FO transformation enjoys proven tight security bound and quantum resistance, but also leaves many open problems. One such problem is the substantial computational cost of using \emph{de-randomization} and \emph{re-encryption} for providing ciphertext integrity. In many post-quantum schemes, including ML-KEM, the input encryption routine is substantially more expensive than the input decryption routine, so running the encryption as a subroutine substantially increases the runtime cost of the decapsulation routine. While not the focus of this project, \emph{re-encryption} also introduces risks of side-channel vulnerabilities that may expose the plaintext or the secret key, as demonstrated in \cite{ravi2019generic} and \cite{ueno2022curse}.

% TODO: trivially combining PKE and digital signature is not authenticated encryption?

Achiving indistinguishability under adaptive chosen ciphertext attack (IND-CCA2) is a solved problem in symmetric cryptography. Given an IND-CPA secure symmetric cipher and an existentially unforgeable message authentication code, combining them in a pattern called ``encrypt-then-MAC'' achieves IND-CCA2 security \cite{DBLP:conf/asiacrypt/BellareN00}\cite{DBLP:conf/eurocrypt/CanettiK01}\cite{DBLP:conf/crypto/Krawczyk01}. However, ``encrypt-then-MAC'' cannot be directly applied within the context of public-key cryptography, because the two parties do not have pre-shared symmetric keys. Our main contribution is a novel generic key encapsulation mechanism (KEM) construction that applies the ``encrypt-then-MAC'' pattern by deriving the symmetric key from some public-key encryption scheme's plaintext.
